---
title: "Batch Correction"
output: html_document
vignette: >
  %\VignetteIndexEntry{Batch Correction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
fig_width: 12 
fig_height: 12 
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE
)
```

# Load data
```{r message=FALSE}
library (TBdev)
```

We have processed 6 scRNAseq datasets and stored them under the `data_dir` as
defined below. Each '.Robj' file contains a Seurat object for that dataset.
Each dataset has already been normalised.
```{r}
root <- '.' #in this case, it is the current directory
data_dir <- paste (root, 'data', sep='/')
save_dir <- paste (root, 'result', sep='/')

save_robj <- c('Xiang_R.Robj', 'Liu_R.Robj', 'Blakeley_R.Robj',
               'Petropoulos_R.Robj', 'Yan_R.Robj', 'Zhou_R.Robj')
# load all datasets as a single list of Seurat object
paper_ID <- gsub ('_R.Robj$', '', save_robj)
all_datasets <- load_all_data (save_robj, data_dir, paper_ID = paper_ID)
# This function will also create a metadata column 'paper' for the `paper_ID`
```

# Merge data
The simplest way to integrate the datasets is to merge all expression matrix
together, as implemented in `merge_seurat` function.
```{r}
merged <- merge_seurat (all_datasets, 
                        assays = c('RNA')) #which assay to merge
merged <- run_dim_red (merged, 
                       find_var_features = T, #call `Seurat::VariableFeatures`
                       normalize=F, #no need to normalise the dataset again
                       var_scale=T, #scale the variable features for PCA
                       run_diff_map=F
)

# The cell type labels have been stored under the column 'Type'
# This step is extremely important for the correct order of appearance of cell
# labels in the figure legend and for assigning colors
merged <- clean_metadata (merged, cell_type_col='Type', date_col='date')
merged$cluster <- paste ('C', merged$seurat_clusters, sep='')
```

Inspect the merging process
```{r, fig.width = 12, fig.height = 12}
AP <- list (pointsize=3, legend_point_size=3, fontsize=11, point_fontsize=4,
            font_fam= 'sans')
plot_dim_red (merged, group.by = c('Type', 'date', 'paper', 'cluster'), 
              DR='pca', AP=AP, further_repel=F)
```
In this case, the datasets seem to be merged well, as the two pre-implantation
dataset (Petropoulos and Yan) are merged at a similar position. Likewise, Xiang
and Zhou, two datasets on pre- and post- implantation trophoblasts, are merged
together. The Liu dataset does not merge with others, probably because it
contains first and second trimester placental samples.

```{r, fig.width = 12, fig.height = 12}
plot_dim_red (merged, group.by = c('Type', 'date', 'paper', 'cluster'), 
              DR='pca', dims=1:3, AP=AP, further_repel=F, 
              all_theta=120, #degree of azimuthal rotation
              vert_just=0.9) #low the axis labels a bit
```

However, batch effects seem to be present from UMAP
```{r, fig.width = 12, fig.height = 12}
plot_dim_red (merged, group.by = c('Type', 'date', 'paper', 'cluster'), 
              DR='umap', AP=AP, further_repel=F)
```

```{r, include=FALSE}
# reduce memory load
rm (merged)
```

# Batch correction
Even though simple merging is successful in this dataset, let us investigate
whether batch correction implementd in Seurat improves the integration process.
Because Blakeley dataset contains fewer than 30 cells, which is lower than
Seurat requirement, the dataset is removed here.

```{r, message=FALSE}
integrated <- batch_correct (all_datasets[1:6!=3], num_anchor_genes=1000)
```

```{r, include=FALSE}
rm (all_datasets)
```

Clean up the metadata
```{r}
integrated <- clean_metadata (integrated, cell_type_col='Type', date_col='date')
integrated$cluster <- paste ('C', integrated$seurat_clusters, sep='')
integrated <- run_dim_red (integrated, assay='integrated', find_var_features =
                           T, normalize=F, var_scale=T, run_diff_map=F)
```

Note that the batch corrected data are in the 'integrated' assay
```{r, fig.width = 12, fig.height = 12}
plot_dim_red (integrated,  
              group.by = c('Type', 'date', 'paper', 'cluster'), 
              DR='pca', AP=AP, further_repel=F, 
              move_y= 1, move_x=1) #shift the arrow axis to avoid view obstrucion
```

The batch effect seems to have been reduced significantly in UMAP. The
UMAP matches well with lineage progression.
```{r, fig.width = 12, fig.height = 12}
plot_dim_red (integrated, group.by = c('Type', 'date', 'paper', 'cluster'), 
              DR='umap', AP=AP, further_repel=F)
```

However, two key problems remain: the stromal cells (STR) overlap with
syncytiotrophoblast (STB) and Oocyte and 4C overlap with cytotrophoblast (CTB).
This is not because they are separated along the third dimension:
```{r, fig.width = 12, fig.height = 12}
integrated <- Seurat::RunUMAP (integrated, assay='integrated', dims=1:30, n.components=3)
plot_dim_red (integrated, group.by = c('Type', 'date', 'paper', 'cluster'), 
              DR='umap', dims=1:3, AP=AP, further_repel=F, all_theta=20)
```
Based on the above reasons, in the paper accompanying this package, we did not
use the Seurat batch correction procedure.
