---
title: "GPLVM"
output: html_document
vignette: >
  %\VignetteIndexEntry{GPLVM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE
)
```

Set up python environment to call the script for GPLVM.
Because the script for GPLVM is written in python, to call this script in R, it
is necessary to install the reticulate package. Next, a python virtual
environment needs to be created as followed:

```{bash, eval= FALSE}
conda create -n ptime python=3.6
conda activate ptime
conda install -c conda-forge tensorflow==1.13.1
conda install pip
pip install gpflow==1.0.0
conda install -c bioconda anndata
```

```{r message = FALSE}
reticulate::use_condaenv("ptime")
# where the python files are located within the package:
script_root <- system.file("python/", package='TBdev')
reticulate::source_python (paste (script_root, 'GPLVM/GrandPrix.py', sep='/'))
reticulate::source_python (paste (script_root, 'STREAM/fitting_prob.py', sep='/'))

library (TBdev)
```

# Load data
Load a seurat object that has been saved in the directory `data`.
```{r, fig.width=12, fig.height=6}
root <- '.'
data_dir <- paste (root, 'data', sep='/')
all_data <- get (load (paste (data_dir, 'final_merged_tb.Robj', sep='/') ))
AP <- list (pointsize=3, legend_point_size=3, fontsize=11, point_fontsize=4,
            font_fam= 'sans')
plot_dim_red (all_data, group.by = c('broad_type', 'date'), DR='pca',
              dims=c(1,2,3), all_theta=50, AP=AP
)
```

Select the cells in the trophoblast lineage
```{r}
# save from cMor
data (CT)
tb_data <- all_data [, ! (all_data$revised %in% c(CT$in_vitro_cells, CT$non_emb_lineage,
                   'Oocyte', 'Zy', '2C', '4C', '8C', 'PE', 'EPI', 'PSA-EPI' ))]
exp_mat <- save_to_csv (tb_data, select_genes=4000)
# scale the data as required by GPLVM
Y <- scale(t(exp_mat))
```

# Embedding
It is highly recommended to run the python script on a GPU. 
That is, save the expression matrix and then run `inst/python/GPLVM/gplvm_tb.py`.
However, for illustration in this R markdown, the code below is run on a CPU
for 100 epochs of iteration only. For this dataset, convergence is usually
reached after 50~100 epochs.
It may be better to set a higher number of epochs e.g. 1000 to ensure
convergence for other datasets in general.
```{r, message = FALSE}
reticulate::py_set_seed (100)
Grand_out = fit_model (data=Y, # rows are cells and columns are genes
                       n_latent_dims=3L, #here 3 latent dimensions are computed
                       n_inducing_points=10L, #number of inducing points
                       latent_prior_var=1., 
                       maxitr=100L, #maximum epoch of iteration
                       disp=T,  return_pred=T,
                       kernel=list ('name'='RBF', 'ls'=1.0, 'var'=1.0)
                       # Use RBF kernel
)
```

```{r include=FALSE}
rm (Y)
```

Visualise the result
```{r, fig.width=12, fig.height=6}
tb_data <- RunGPLVM (tb_data, Grand_out[[1]][[1]])
plot_dim_red (tb_data, group.by = c('broad_type', 'date'), DR='gplvm',
              dims=c(1,2,3), all_theta=50, AP=AP
)
```

Another benefit of GPLVM is infer the gene expression distribution at given
pseudotime location. We can determine the probability of a cell whose
expression fits into this distribution.
To illustrate, we will compare the human trophoblast stem cells (hTSC) with
those along the trophoblast lineage. Human embryonic stem cells (hESC) will
also be compared.

```{r}
in_vitro <- all_data [rownames (exp_mat),all_data$broad_type %in% c('hTSC_OKAE', 
                                                        'hTSC_TURCO', 'hESC')]
in_vitro_mat <- as.matrix (Seurat::GetAssayData (in_vitro, assay='RNA', slot='data'))
in_vitro_met <- in_vitro@meta.data
fit_chance <- fitting_likelihood (data.frame (exp_mat), 
                                  new_X = data.frame (in_vitro_mat), 
                                  new_meta = in_vitro_met, 
                                  group_by = 'broad_type',
                                  pt_mean = data.frame (Grand_out[[2]][[1]]),
                                  pt_var  = data.frame (Grand_out[[2]][[2]])
)
```

Visualise the results on the latent dimension inferred by GPLVM
```{r, fig.width=12, fig.height=10}
meta <- tb_data@meta.data
tb_data@meta.data <- cbind (meta [, !colnames (meta) %in% colnames (fit_chance)], fit_chance)
# shorten the name a bit for plotting
colnames (tb_data@meta.data) <- gsub ('^hTSC_', '', colnames (tb_data@meta.data))
plot_dim_red (tb_data, group.by = c('broad_type', 'hESC', 'OKAE', 'TURCO'), 
              DR='gplvm', dims=c(1,2,3), all_theta=50, AP=AP, 
              label_col='broad_type' #label the cell types in all subplots
) 
```

```{r include=FALSE}
rm (list = ls ())
```
